---
title: "PredictionAssignmentWriteup"
author: "Sophie"
date: "8/9/2021"
output:
  pdf_document: default
  html_document: default
---
1.Overview
The main goal of the project is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and predict the manner in which they did the exercise. The machine learning algorithm is applied to build the best model and apply it to 20 different test cases.

loading the packages

```{r}
library(caret)
library(lattice)
library(ggplot2)
```


2.Loading and observing the data.

```{r}
trainingdata <- read.csv("~/pml-training.csv")
testingdata <- read.csv("~/pml-testing.csv")
dim(trainingdata)
dim(testingdata)
```

3.Cleaning up the data and remove the irrelevant variables and missing values.

```{r}
trainingdata <- trainingdata[,-(1:7)]
NZV <- nearZeroVar(trainingdata)
trainingdata <- trainingdata[,-NZV]
PNA<- sapply(trainingdata,function(x)mean(is.na(x)))>0.95
trainingdata <- trainingdata[,PNA==FALSE]
```

4.Creating training and testing set.

```{r}
set.seed(333)
inTrain <- createDataPartition(trainingdata$classe,p=0.6,list=FALSE)
training <- trainingdata[inTrain,]
testing <- trainingdata[-inTrain,]
dim(training)
dim(testing)
```

5.Building 3 models

```{r}
modrpart <- train(classe~.,data=training,method="rpart")
modRF <- train(classe~.,data=training,method="rf")
modGBM <- train(classe~.,data=training,method="gbm",verbose = FALSE)
```

6.Assessing these models and find the most accurate model.

```{r}
testing$classe <- as.factor(testing$classe)
predrpart <- predict(modrpart,newdata=testing)
cmrpart <- confusionMatrix(predrpart,testing$classe)
predRF <- predict(modRF,newdata=testing)
cmRF <- confusionMatrix(predRF,testing$classe)
predGBM <- predict(modGBM,newdata=testing)
cmGBM <- confusionMatrix(predGBM,testing$classe)
cmrpart$overall[1]
cmRF$overall[1]
cmGBM$overall[1]
```

7.Predicting the test data with the best model RF.

```{r}
pred <- predict(modRF,testingdata)
print(pred)
```

